---
title: "segs_redo"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

## sample gfw_data to csv

Ran on [rstudio.ships4whales.org](http://rstudio.ships4whales.org) to produce 3 days worth of data from points `gfw_data`: `gfw_data_2020-04-02-to-04.csv`.

```{python, eval=F}
import pandas as pd
from google.cloud import bigquery
from google.oauth2 import service_account
import csv

#FILL IN YOUR PATH TO THE 'Benioff Ocean Initiative-454f666d1896.json'
credentials_json = '/home/admin/Benioff Ocean Initiative-454f666d1896.json'
credentials = service_account.Credentials.from_service_account_file(credentials_json)

project_id = 'benioff-ocean-initiative'
client = bigquery.Client(credentials= credentials,project=project_id)

sql = "SELECT MAX(timestamp) AS timestamp_max FROM clustered_datasets.gfw_data WHERE date(timestamp) > '2020-01-01';"
client.query(sql).to_dataframe()
# timestamp_max = 2020-04-04 23:59:58+00:00

# get 3 days of data (timestamp_max = 2020-04-04 23:59:58+00:00)
sql = """
  SELECT
    mmsi, timestamp, lon, lat, speed_knots, implied_speed_knots, source
  FROM
    `benioff-ocean-initiative.clustered_datasets.gfw_data`
  WHERE 
    DATE(timestamp) > '2020-04-01'
  """
df = client.query(sql).to_dataframe()
df.to_csv('/home/admin/plumber-api/gfw_data_2020-04-02-to-04.csv')
```

## get top3 ships

```{r}
library(tidyverse)
library(here)
library(sf)
library(units)
library(lubridate)
library(mapview)
library(leaflet)
library(DT)
# library(bigrquery)
# library(rmapshaper)
library(fs)
library(leafsync)
here = here::here

d_csv       <- here("data/gfw_data_2020-04-02-to-04.csv")
top3_csv    <- here("data/gfw_data_2020-04-02-to-04_top3.csv")
lns_geo     <- here("data/gfw_data_2020-04-02-to-04_top3_lns.geojson")
# lns_s05_geo <- here("data/gfw_data_2020-04-02-to-04_top3_lns_s05.geojson")
# lns_s3_geo  <- here("data/gfw_data_2020-04-02-to-04_top3_lns_s3.geojson")
lns_d01_geo <- here("data/gfw_data_2020-04-02-to-04_top3_lns_d01.geojson")
lns_d05_geo <- here("data/gfw_data_2020-04-02-to-04_top3_lns_d05.geojson")
lns_d1_geo  <- here("data/gfw_data_2020-04-02-to-04_top3_lns_d1.geojson")

#service_account_json <- "/home/admin/Benioff Ocean Initiative-454f666d1896.json"
# service_account_json <- "/Volumes/GoogleDrive/My Drive/projects/whalesafe/data/gfw/Benioff Ocean Initiative-454f666d1896.json"
# 
# bq_auth(path = service_account_json)
# bq <- dbConnect(
#   bigquery(),
#   project = "benioff-ocean-initiative",
#   dataset = "whalesafe_ais",
#   billing = "benioff-ocean-initiative")
```


```{r}
d <- read_csv(d_csv) %>% 
  arrange(mmsi, timestamp)
d %>% head() %>% datatable()

d_n <- d %>% 
  group_by(mmsi) %>% 
  summarise(
    n = n(),
    implied_speed_knots_avg = mean(implied_speed_knots, na.rm=T)) %>% 
  ungroup() %>% 
  arrange(desc(n))
d_n %>% datatable()

top3 <- d_n %>% 
  slice(1:100) %>% 
  arrange(desc(implied_speed_knots_avg)) %>% 
  slice(1:3)

d %>% 
  filter(mmsi %in% top3$mmsi) %>% 
  write_csv(top3_csv)

top3 %>% datatable()
```


## calculate speed bins

TODO: Add last row of ships already in from previous day if present at beginning.

```{r}
pts <- read_csv(top3_csv) %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326, remove = F)
mapview(pts)

empty <- st_as_sfc("POINT(EMPTY)")
pts <- pts %>%
  arrange(mmsi, timestamp) %>% 
  mutate(
    date = date(timestamp)) %>% 
  group_by(mmsi) %>% # IMPORTANT!
  mutate(
    calc_time_s      = difftime(
      lead(timestamp), timestamp, units = "secs") %>% 
      as.numeric(),
    calc_dist_m      = st_distance(
      geometry, lead(geometry, default = empty), by_element = TRUE) %>%
      set_units(m) %>% as.numeric(),
    speed_calc_knots = calc_dist_m / calc_time_s * 1.94384)

hist(pts$speed_calc_knots)

speed_bins = c(0, 10, 12, 15, Inf)
pts <- pts %>%
  mutate(
    speed_calc_bin = cut(speed_calc_knots, speed_bins, include.lowest = T))

ggplot(aes(x = speed_calc_bin), data = pts) +
  geom_bar()
```

## differentiate segments

```{r}
# identify rows ending a segment
pts <- pts %>%
  mutate(
    seg_chg = (lead(speed_calc_bin) != speed_calc_bin), 
    # set NA endpoints to 0 to include in segment
    seg_chg = replace_na(seg_chg, F)) %>% 
  ungroup()

# repeat rows to begin the next segment
pts <- rbind(
  # original points
  pts,
  # duplicate points for change in speed
  pts %>% 
    filter(
      seg_chg) %>% 
    mutate(
      seg_chg = F),
  # duplicate points for previous day
  pts %>% 
    group_by(mmsi, date) %>% 
    filter(
      timestamp == max(timestamp)) %>% 
    mutate(
      seg_chg = F) %>% 
    # drop last day though
    group_by(mmsi) %>% 
    filter(
      timestamp != max(timestamp)) %>% 
    # add row to next day
    ungroup() %>%
    mutate(
      date = date + days(1))) %>% 
  # arrange and group
  arrange(
    mmsi, timestamp, seg_chg) %>% 
  group_by(mmsi, date)

# create unique segment id
pts <- pts %>%
  mutate(
    seg_id  = cumsum(seg_chg))

# check that every row has a seg_id assigned
stopifnot(sum(is.na(pts$seg_id)) == 0) 
```

## construct line segments

```{r}
lns <- pts %>%
  group_by(mmsi, date, seg_id) %>% 
  summarize(
    speed_calc_bin       = first(speed_calc_bin),
    timestamp_beg        = min(timestamp),
    timestamp_end        = max(timestamp),
    speed_calc_knots_avg = mean(speed_calc_knots, na.rm = T),
    geom = st_union(geometry) %>% 
      st_cast("LINESTRING")) %>% 
  # remove lines with < 2 points
  mutate(
    npts = map_int(geom, function(x) npts(x))) %>% 
  filter(npts > 1)

write_sf(lns, lns_geo, delete_dsn = T)
```

## map segments

```{r}
lns <- read_sf(lns_geo)

speed_calc_bins <- c("[ 0, 10 ]", "(10,12]", "(12,15]", "(15,Inf]")

stopifnot(length(setdiff(unique(lns$speed_calc_bin), speed_calc_bins)) == 0)

lns$speed_calc_bin <- factor(lns$speed_calc_bin, speed_calc_bins, ordered = T)

pal <- colorFactor("YlOrRd", lns$speed_calc_bin, ordered = T)

map_lns <- leaflet() %>%
  addProviderTiles(
    providers$Esri.OceanBasemap,
    options = providerTileOptions(
      opacity = 0.7)) %>% 
  addPolylines(
    data = lns %>% filter(!is.na(seg_id)),
    color = ~pal(speed_calc_bin),
    label = ~speed_calc_bin,
    opacity = 0.7) %>% 
  addLegend(
    data = lns,
    pal = pal, values = ~speed_calc_bin, title = "Speed bin (knots)")
map_lns
```

## simplify line segments

OLD: `rmapshaper::ms_simplify()`

```{r, eval = F}
# keep 5% of points (default)
lns_s05 <- ms_simplify(lns, keep = 0.05) %>% 
  mutate(
    npts = map_int(geometry, function(x) npts(x)))
write_sf(lns_s05, lns_s05_geo, delete_dsn = T)

# keep 30% of points
lns_s3 <- ms_simplify(lns, keep = 0.3) %>% 
  mutate(
    npts = map_int(geometry, function(x) npts(x)))
write_sf(lns_s3, lns_s3_geo, delete_dsn = T)
```

NEW: `sf::st_simplify()` more similar to PostGIS `ST_Simplify()`

```{r}
lns_smpl <- lns %>% 
  # simplify and keep track of number of points per segment
  mutate(
    geom_d01    = st_simplify(geometry, dTolerance = 0.01),
    geom_d05    = st_simplify(geometry, dTolerance = 0.05),
    geom_d1     = st_simplify(geometry, dTolerance = 0.1),
    npts_d01    = map_int(geom_d01, function(x) npts(x)),
    npts_d05    = map_int(geom_d05, function(x) npts(x)),
    npts_d1     = map_int(geom_d1 , function(x) npts(x)),
    npts_d01dif = npts - npts_d01,
    npts_d05dif = npts - npts_d05,
    npts_d1dif  = npts - npts_d1) %>% 
  arrange(desc(npts_d1dif))

lns_d <- lns_smpl %>% 
  st_drop_geometry() %>% 
  select(
    mmsi, date, 
    timestamp_beg, timestamp_end,
    speed_calc_bin, speed_calc_knots_avg,
    starts_with("npts"))
datatable(lns_d)

lns_d %>% 
  st_set_geometry(lns_smpl$geom_d01) %>% 
  write_sf(lns_d01_geo, delete_dsn = T)
lns_d %>% 
  st_set_geometry(lns_smpl$geom_d05) %>% 
  write_sf(lns_d05_geo, delete_dsn = T)
lns_d %>% 
  st_set_geometry(lns_smpl$geom_d1) %>% 
  write_sf(lns_d1_geo, delete_dsn = T)

# display file sizes
dir_info(here("data"), regexp = "(top3_lns\\.geojson$)|(lns_d[0-9]+\\.geojson$)") %>% 
  mutate(
    file = basename(path)) %>% 
  select(file, size)
```

## map compare simplified segments

```{r}
lns_d05 <- read_sf(lns_d05_geo)
lns_d05$speed_calc_bin <- factor(lns_d05$speed_calc_bin, speed_calc_bins, ordered = T)

map_lns_d05 <- leaflet(lns_d05) %>%
  addProviderTiles(
    providers$Esri.OceanBasemap,
    options = providerTileOptions(
      opacity = 0.7)) %>% 
  addPolylines(
    color = ~pal(speed_calc_bin),
    label = ~speed_calc_bin,
    opacity = 0.7) %>% 
  addLegend(
    pal = pal, values = ~speed_calc_bin, title = "Speed bin (knots)<br> <small>st_simplify(dTolerance=0.05)</small>")

sync(map_lns, map_lns_d05)
```

